---
title: "Data Science for Public Policy"
subtitle: "Assignment 04"
author: "Juan Menendez - jmm676 & Diana Rivas - dnr36"
execute:
  warning: false
format:
  html:
    embed-resources: true
---

## Data: Latin American countries in the 2000's

This data set contains selected indicators from census data at the individual level for 38 Latin American countries for the 2000-2010 decade. Source: IPUMS International ([Codebook](https://live.ipumsi.datadownload.ipums.org/web/extracts/ipumsi/2210604/ipumsi_00009.xml?jwt=eyJhbGciOiJSUzI1NiJ9.eyJpc3MiOiJ1bWEiLCJleHAiOjE3MDg2NjIwMzksIm5iZiI6MTcwODY2MTEzOSwic3ViIjoiN2RiMTViMTAtZjFjYi0wMTM1LWFmMmUtMDA1MDU2YTM1NDA1IiwiY29sbGVjdGlvbnMiOlsiY3BzIiwiaXB1bXNpIl19.D8QH7AA7EZluCH9dSzLSjOph6c6SD6UqifJGCQdwtB70yb-56nkF_ZrhqeAc-M2KubzNl7BV-4qp3YVq08znUFfm2waK6pWmOiGFFE2MRYAgLkfnwZniPCvRbuesOwy8Omv7XnMx217Th7aGhLWgfZkq9FU8mCq-7N1VVzpWZk2t6_SAf0mzRrfbD6CsnDNNWLDu2IE85mmr38zOgDEBDwqL5kIsuV_HKuFzFAXx0clxkmf7oEb1S12tftE0v7H-UQtEBgq_nRUU3K7fBK2tqnDCVeTNRUWXJ8UPcpIRZCzmqyZ9plt9Qt3zmux9QxRuWwY11jts4DOzFQfNZwxtHw#SAMPLE))



## Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
